---
title: "Forecasting"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---
TASK 1

Introduction
In this we will be analysing potential effects of both climate and pollution on disease specific mortality between the years 2010-2020 and forecasting in terms of n terms of R squared, AIC, BIC, MASE et by building different forecasting models such as DLM, ARDL, polyck, koyck, dynamic, exponential smoothing and state-space model.



Data Description
This dataset contains potential effects of both climate and pollution on disease specific mortality between the years 2010-2020

```{r echo = TRUE}
mort<-read.csv("C:/Users/glane/Downloads/MORT.csv")
head(mort)
class(mort)
```

Data Preparation
We will be converting the dataset into time series inorder to proceed with further analysis and exploration.

```{r}
mortality <- ts(mort$mortality, frequency = 52 ,start = 2010)
temp <- ts(mort$temp, frequency = 52 ,start = 2010)
```
Plot the data 
```{r}
plot(mortality, main = "Time series for monthly average mortality")
plot(temp, main = "Time series for monthly average temperature")
```
From this we can see a few things

1.The mortality peaked only once that post 2012 and was 30.43, apart from that it's either till 20 or below that
2.The trend is not constant, it moves down during the recession
3.There is differences in mortality based on month

Temperature plot
```{r}
plot(temp, main = "Time series for monthly average temperature")
```
There is a clear trend where temperature peaks during early months of the year


ACF Plot and ADF Test for mortality
```{r}
acf(mortality, lag.max = 48, main="Sample ACF for Monthly average mortality",
cex.main = 0.65)
```
```{r}
library(urca)
summary(ur.df(mortality, type = "none", lags = 1, selectlags = "AIC"))
```
Analysing the ACF Plot andAugmented Dickey-Fuller test results for monthly mortality we can see that,
There is an obvious existence of seasonality in the ACF plot
ADF Test gives a p value less than 0.05, where we can reject the null hypothesis of non-stationarity


ACF Plot and ADF Test for temperature
```{r}
acf(temp, lag.max = 48, main="Sample ACF for Monthly average temp",
cex.main = 0.65)
```
```{r}
summary(ur.df(temp, type = "none", lags = 1, selectlags = "AIC"))
```
Analysing the ACF Plot andAugmented Dickey-Fuller test results for monthly mortality we can see that,
There is an obvious existence of seasonality in the ACF plot
ADF Test gives a p value less than 0.05, where we can reject the null hypothesis of non-stationarity



Time Series Decomposition
You should decompose your data to get or present a basic understanding of your data.
Outputs of a decomposition

1.The underlying trend of your data
2.A seasonal factor
3.A remainder which explains what the trend and seasonal factor do not

Types of seasonal decompositions

Additive Seasonal Decomposition: Each season gets moved by a constant number that is added or subtracted from the trend.
Multiplicative Seasonal Decomposition: Each season has a number we multiply to the trend.

```{r}
decomposed_mortality_multi <- decompose(mortality, type = "additive")
plot(decomposed_mortality_multi)

decomposed_temp_additive <- decompose(temp, type = "additive")
plot(decomposed_temp_additive)

```

Correlation Analysis
```{r}
cor(mort)
```
Model Fitting
Distributed Lag Models
```{r}
library(dLagM)
library(forecast)
model.dlm = dlm(x = mort$mortality, y = mort$temp ,
q = 4 )
summary(model.dlm)
AIC(model.dlm)
BIC(model.dlm)
MASE(model.dlm)
checkresiduals(model.dlm$model, test = F)
```

Polynomial Distributed Lag Models
```{r}
model.poly = polyDlm(x = mort$mortality, y = mort$temp ,
q = 4 , k = 2 , show.beta = TRUE)
summary(model.poly)
AIC(model.poly)
BIC(model.poly)
MASE(model.poly)
checkresiduals(model.poly$model, test = F)
```
Koyck Model Models
```{r}

model.koyck = koyckDlm(x = mort$mortality, y = mort$temp)
summary(model.koyck )
AIC(model.koyck )
BIC(model.koyck )
MASE(model.koyck )
checkresiduals(model.koyck$model, test = F)
```
autoregressive DLM
```{r}
model.ardl =  ardlDlm(x = mort$mortality, y = mort$temp,p =1,q = 1)
summary(model.ardl )
AIC(model.ardl )
BIC(model.ardl )
MASE(model.ardl )
checkresiduals(model.ardl$model, test = F)
```

```{r}
library(dynlm)
model.dynlm =  dynlm(mortality~ L(mortality , k = 1 ) + season(mortality))
summary(model.dynlm )
AIC(model.dynlm )
BIC(model.dynlm )
accuracy(model.dynlm)   #For MASE
residuals(model.dynlm$model)
```

Exponential Smoothing Models
Holt-Winters Additive Method
```{r}
mortality1 <- ts(mort$mortality, frequency = 20 ,start = 2010)
model.hw <- hw(mortality1,seasonal = "additive" , h = 2*frequency(mortality1))
accuracy(model.hw)
checkresiduals(model.hw$model)
```

Holt-Winters Additive Method - Damped

```{r}
model.hwd <- hw(mortality1,seasonal = "additive" , h = 2*frequency(mortality1),damped = T)
accuracy(model.hwd)  
checkresiduals(model.hwd$model)
```


Holt-Winters Multiplicative Method
```{r}
model.hwm <- hw(mortality1,seasonal = "multiplicative" , h = 2*frequency(mortality1))
accuracy(model.hwm)
checkresiduals(model.hwm$model)
```


Holt-Winters Multiplicative Method - Damped
```{r}
model.hwmd <- hw(mortality1,seasonal = "multiplicative" , h = 2*frequency(mortality1))
accuracy(model.hwmd)
checkresiduals(model.hwmd$model)
```

State Space Models
```{r}
model.ets = ets(mortality1, model="ZZZ")
summary(model.ets)
accuracy(model.ets)
checkresiduals(model.ets)
```

```{r}
model_data_df = data.frame(Model=character(),MASE=numeric(),
                           BIC= numeric(),AIC=numeric())

model_data_df = rbind(model_data_df,cbind(Model="DLM",MASE=MASE(model.dlm),
                                               BIC = BIC(model.dlm),
                                               AIC = AIC(model.dlm)))

model_data_df = rbind(model_data_df,cbind(Model="polyDlm",MASE=MASE(model.poly),
                                               BIC = BIC(model.dlm),
                                               AIC = AIC(model.dlm)))
                      
model_data_df = rbind(model_data_df,cbind(Model="koyckDlm",MASE=MASE(model.koyck),
                                               BIC = BIC(model.koyck),
                                               AIC = AIC(model.koyck)))

model_data_df = rbind(model_data_df,cbind(Model="ARDlm",MASE=MASE(model.ardl),
                                               BIC = BIC(model.ardl),
                                               AIC = AIC(model.ardl)))

model_data_df = rbind(model_data_df,cbind(Model="DYNLM",MASE=accuracy(model.dynlm)[6],
                                               BIC = BIC(model.dynlm),
                                               AIC = AIC(model.dynlm)))

model_data_df = rbind(model_data_df,cbind(Model="HWA",MASE= accuracy(model.hw)[6],
                                               BIC = model.hw$model$bic,
                                               AIC = model.hw$model$aic))

model_data_df = rbind(model_data_df,cbind(Model="HWAD",MASE=accuracy(model.hwd)[6],
                                               BIC = model.hwd$model$bic,
                                               AIC = model.hwd$model$aic))

model_data_df = rbind(model_data_df,cbind(Model="HWM",MASE=accuracy(model.hwm)[6],
                                               BIC = model.hwm$model$bic,
                                               AIC = model.hwm$model$aic))

model_data_df = rbind(model_data_df,cbind(Model="HWMD",MASE=accuracy(model.hwmd)[6],
                                               BIC = model.hwmd$model$bic,
                                               AIC = model.hwmd$model$aic))

model_data_df = rbind(model_data_df,cbind(Model="ETS",MASE=accuracy(model.ets)[6],
                                               BIC = model.ets$bic,
                                               AIC = model.ets$aic))
```
```{r}
knitr::kable(model_data_df, caption = "Model Parameters captured so far")
```
Holt - winter's additve without damped has the lowest MASE and it is the optimal model. 

```{r}
predict(model.hw,n.ahead = 4)
futurVal <- forecast(model.hw,h=4)
plot(futurVal)
```